An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-10-09 aistats2017.tex
Removed several tiny and scriptsize commands. I'm almost certain reviewers will hate that

%\begin{figure}[t]
% \centering
%\includegraphics[width=1\textwidth]{mh_convolution_diagram_v2}
%    \caption{
%   Three examples of $X_{\rm norm}$ and $X_{\rm corr}$ distributions that
%    convolve to form the standard logistic distribution. These were obtained
%    based on our original linear programming formulation, as discussed in
%    Section~\ref{sec:correction}. We use three standard deviation values of
%    $X_{\rm norm}$. The two red curves convolve to form the logistic, etc. The
%    $y$-axis is capped at 0.5 for readability. This figure must be viewed in
%    color.
%    }
%    \label{fig:deconvolution}
%\vspace{-10pt}
%\end{figure}

@@

For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    \small{C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|}
\end{equation}
To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}$ and $j \in \{-N,
\ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$, and the
condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
\begin{equation}\label{eq:conv2}
    \scriptsize{C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \{-2N,\ldots,2N\}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.}
\end{equation}
To make the problem easier to understand, we can define a matrix $M$ and vectors
$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
Equation~(\ref{eq:conv2}) is equivalent to minimizing the $L_{\infty}$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
\end{equation}

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.
%Figure~\ref{fig:deconvolution} shows three examples of $X_{\rm norm}$ and
%$X_{\rm corr}$ densities that sum to $X_{\rm log}$ from this process.

@@

\begin{lemma}\label{lem:quant_clt}
Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
random variables with sample mean $\bar{X}$ and sample variance $s^2_X$
where:
\begin{equation}
    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i \qquad {\rm and} \qquad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
\end{equation}
This implies $t=\bar{X}/s_X$ has an approximate Student's distribution
which approaches a normal distribution in the limit. Then
\begin{equation}\label{eq:clt-bounds}
    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
\end{equation}
\end{lemma}

\begin{proof}
The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
\begin{equation}\label{eq:clt-boundsx}
\scriptsize{-6.4E|X|^3-2E|X| \leq    \sup_x|{\rm Pr}(t<x) - \Phi(x)|{\sqrt{n}} \leq 1.36E|X|^3}
\end{equation}
This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
and combining with $x>0$, we obtain the weaker but unqualified bound
in Equation~(\ref{eq:clt-bounds}).
\end{proof}

@@

\begin{proof}
We first observe that
\begin{equation}\label{eq:pdf_difference}
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\end{equation}
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that for all $z$:
\begin{equation}
\begin{split}
\tiny{-\epsilon} & = \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
& \tiny{\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx }\\
& \tiny{\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon}\\
\end{split}
\end{equation}
\end{proof}

@@

Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we observe that:

@@

\begin{center}
{\Large Outline of Appendix}
\end{center}

In this appendix, we describe the following topics:

\begin{itemize}[noitemsep]
    \item The proof of Lemma~\ref{lem:worst_case}.
\end{itemize}

@@

%% \section{Gaussian Mixture Experiment Details}\label{app:gaussian}

%% In this section, we provide more details on Section~\ref{ssec:gaussians}.  Our
%% parameter is a 2-D vector
%% $\theta = \langle \theta_1, \theta_2 \rangle$, where
%% \begin{equation}
%%     \theta_1 \sim \mathcal{N}(0, \sigma_1^2) \quad \mbox{ and } \quad \theta_2 \sim \mathcal{N}(0, \sigma_2^2)
%% \end{equation}
%% and where $\mathcal{N}$ indicates the normal distribution (more generally, the
%% multivariate normal). We consider the above as our prior.
%% Following~\cite{langevin_2011}, we set $\sigma_1^2 = 10$ and $\sigma_2^2=1$, so
%% the covariance matrix of $\theta$ is $\Sigma = {\rm diag}(10,1)$. Therefore, the
%% log prior probability we endow on $\theta$ is
%% \begin{equation}
%%     \log p(\theta) = \log \left(\frac{1}{2\pi\sqrt{10}}\right) - \frac{1}{2}\theta^T\Sigma^{-1}\theta.
%% \end{equation}
%% To generate the data, we use the following Gaussian mixture with tied means:
%% \begin{equation}\label{eq:x_points}
%%     x_i \sim \frac{1}{2}\mathcal{N}(\theta_1, \sigma_x^2) + \frac{1}{2}\mathcal{N}(\theta_1+\theta_2, \sigma_x^2)
%% \end{equation}
%% where, again following~\cite{langevin_2011}, we set $\sigma_x^2 = 2$ and we draw
%% the data with $\theta_1=0$ and $\theta_2=1$ fixed (though these values are
%% arbitrary, and different choices would yield different posterior shapes).
%% This means the log likelihood of a single data instance is
%% \begin{equation}\label{eq:log_ll_single_instance}
%%     \log p(x_i \mid \theta) = \log \frac{1}{4\sqrt{\pi}} +
%%     \log\left(\exp\left(-\frac{1}{4}(x_i - \theta_1)^2\right) + \exp\left(-\frac{1}{4}(x_i -
%%     (\theta_1+\theta_2))^2\right)\right)
%% \end{equation}
%% The problem statement is as follows: given some number of conditionally
%% independent data points $x_1, x_2, \ldots, x_N$ generated according
%% to~(\ref{eq:x_points}), determine the posterior distribution of $\theta$:
%% \begin{equation}\label{eq:log_post}
%%     \log p(\theta \mid x_1,\ldots,x_N) = \log p(\theta) + \sum_{i=1}^N\log p(x_i \mid \theta).
%% \end{equation}
%% Alternatively, depending on the application, we may opt to instead pick a point
%% estimate\footnote{We do this for the logistic regression experiment, described
%% in Section~\ref{ssec:logistic}.} of $\theta$, such as the MAP estimate. (If $N$
%% is extremely large, it will cause the posterior to peak sharply at its modes,
%% reducing distribution estimates to point estimates.)

%% One technique we use to make our test more applicable is to add a temperature to
%% the posterior. In general, we will add $T > 1$ so that the posterior is
%% $p(\theta)(\prod_{i=1}^N p(x_i\mid \theta))^{1/T}$, resulting in the log
%% posterior of 
%% \begin{equation}\label{eq:log_prior_temp}
%%     \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\sum_{i=1}^N\log p(x_i \mid \theta).
%% \end{equation}
%% which has the extra $1/T$ to decrease the scale factor, so the variance scales
%% as $1/T^2$.

% Daniel: I decided that I don't like this figure anymore.
% \begin{figure}[t]
%   \centering
%   \includegraphics[width=1\linewidth]{contour_v1}
%   \caption{The posterior distribution, from 50 to 100k samples, with temperature set at 1.}
%   \label{fig:contour1}
% \end{figure}

%% \begin{figure}[t]
%%   \centering
%%   \includegraphics[width=1\linewidth]{contour_v2}
%%   \caption{The posterior distribution, with $N=10000$ and temperature $T$ varying.}
%%   \label{fig:contour2}
%% \end{figure}

%% To gain intuition on the shape of the posterior, Figure~\ref{fig:contour2} shows
%% simulated contour plots of the posterior based on $N=10000$ data points, and
%% temperature varying.  The (non-tempered) posterior is a multimodal distribution
%% with modes at $(0,1)$ and $(1,-1)$. A larger $T$ implies a flatter posterior,
%% which can be useful for parameter exploration so that the samples do not get
%% trapped in a mode.

% Daniel: This was some stuff from the Logistic Regression appendix, but we no
% longer need that.

%% In the logistic regression experiment of Section~\ref{ssec:logistic}, the
%% feature vector for an image consists of its pixel values, normalized between 0
%% and 1. For simplicity, we consider the binary classification case, so we only
%% use digits one ($y = 1$) and seven ($y = -1$).  The probability of the $i^{\rm
%% th}$ output $y_i \in \{-1,+1\}$ with the input vector $x_i$ is
%% \begin{equation}\label{eq:logistic_probability}
%% p(y_i \mid x_i) = \frac{1}{1+\exp (-y_i \theta^T x_i)},
%% \end{equation}
%% where $\theta$ is the 784-length parameter vector.

% Daniel: here are example LaTeX codes for figures and tables if we want to use them.
%\begin{figure}[h]
%  \centering
%  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
%  \caption{Sample figure caption.}
%\end{figure}
%\begin{table}[t]
%  \caption{Sample table title}
%  \label{sample-table}
%  \centering
%  \begin{tabular}{lll}
%    \toprule
%    \multicolumn{2}{c}{Part}                   \\
%    \cmidrule{1-2}
%    Name     & Description     & Size ($\mu$m) \\
%    \midrule
%    Dendrite & Input terminal  & $\sim$100     \\
%    Axon     & Output terminal & $\sim$10      \\
%    Soma     & Cell body       & up to $10^6$  \\
%    \bottomrule
%  \end{tabular}
%\end{table}
%\usepackage[pdftex]{graphicx} ...
%\includegraphics[width=0.8\linewidth]{myfile.pdf}

>>>>

@@

For computation of $C_\sigma$, we assume that its input $Y$ and
another variable $X$ lie in the intervals $[-V,V]$ and $[-2V,2V]$, respectively.
In continuous form, the optimization problem is therefore
\begin{equation}\label{eq:conv1}
    \small{C_\sigma^* = \argmin_{C\sigma} \; \sup_{x \in [-2V,2V]}\left|\int_{-V}^{V}\Phi_{\sigma}(x-y) C_{\sigma}(y)dy - S(x)\right|}
\end{equation}
To get $C_\sigma^*$ in a computable form, we discretize $X$ and $Y$ into $4N+1$
and $2N+1$ values respectively. If $i \in \{-2N, \ldots, 2N\}=\mathcal{I}$ and
$j \in \{-N, \ldots, N\}$, then we can write $X_i = i(V/N)$ and $Y_j = j(V/N)$,
and the condition in Equation~(\ref{eq:conv1}) can be re-expressed as:
\begin{equation}\label{eq:conv2}
    C_\sigma^* = \argmin_{C_\sigma} \max_{i \in \mathcal{I}}\left|\sum_{j = -N}^{N}\Phi_{\sigma}(X_i-Y_j) C_{\sigma}(Y_j) - S(X_i)\right|.
\end{equation}
To make the problem easier to understand, we can define a matrix $M$ and vectors
$u$ and $v$ such that $M_{ij} = \Phi_{\sigma}(X_i-Y_j)$, $u_j =
C_{\sigma}(Y_j)$, and $v_i = S(X_i)$, where the indices $i$ and $j$ are
appropriately translated to be non-negative indices for $M, u,$ and $v$. Thus,
Equation~(\ref{eq:conv2}) is equivalent to minimizing the $L_{\infty}$ norm:
\begin{equation}\label{eq:optimization_l1}
    u^* = \argmin_{u} \|Mu-v\|_{\infty}.
\end{equation}

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $\|Mu - v\|_\infty$ subject to $u \ge 0$.

@@

\begin{lemma}\label{lem:quant_clt}
Let $X_1,\ldots,X_n$ be a set of zero-mean, independent, identically-distributed
random variables with sample mean $\bar{X}$ and sample variance $s^2_X$
where:
\begin{equation}
    \bar{X} = \frac{1}{n}\sum_{i=1}^nX_i, \quad s_X = \frac{1}{n}\left(\sum_{i=1}^n(X_i-\bar{X}^2\right)^{\frac{1}{2}}.
\end{equation}
This implies $t=\bar{X}/s_X$ has an approximate Student's distribution
which approaches a normal distribution in the limit. Then
\begin{equation}\label{eq:clt-bounds}
    \sup_x|{\rm Pr}(t<x) - \Phi(x)| \leq \frac{6.4E|X|^3+2E|X|}{\sqrt{n}}.
\end{equation}
\end{lemma}

\begin{proof}
The following bound is given immediately after Corollary 2 from~\citet{explicit-clt05}:
\begin{equation}\label{eq:clt-boundsx}
\begin{split}
-6.4E|X|^3-2E|X| &\leq \sup_x|{\rm Pr}(t<x)-\Phi(x)|{\sqrt{n}} \\
&\leq 1.36E|X|^3.
\end{split}
\end{equation}
This bound applies to $x\geq 0$. Applying the bound to $-x$ when $x<0$
and combining with $x>0$, we obtain the weaker but unqualified bound
in Equation~(\ref{eq:clt-bounds}).
\end{proof}

@@

\begin{proof}
We first observe that
\begin{equation}\label{eq:pdf_difference}
    P'(z) - Q'(z) = \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx,
\end{equation}
and since $\sup_x|P(x)-Q(x)|\leq \epsilon$ it follows that $\forall z$:
\begin{equation}
\begin{split}
-\epsilon &= \tiny{\int_{-\infty}^{+\infty} -\epsilon R(x) dx } \\
&\leq \int_{-\infty}^{+\infty}(P(z-x)-Q(z-x))R(x) dx \\
&\leq \int_{-\infty}^{+\infty}\epsilon R(x) dx = \epsilon.
\end{split}
\end{equation}
\end{proof}
\vspace{-10pt}

@@

Relating Lemma~\ref{lem:cdf_bounds} to our acceptance test, we have the
following Corollary:

@@

\begin{center}
{\Large Supplementary Material}
\end{center}

@@
